---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

getwd()
locpath= getwd()
setwd(locpath)

library(pacman)
p_load(plyr, dplyr, stringr, tidyverse, tidyr, ggplot2, pastecs, lmerTest, MuMIn, lme4, modelr, Metrics, caret, ddalpha, ggplot2, pastecs, crqa, lmerTest, pROC, modelr) # choose n (y/n?) in ddalpha
library(rethinking)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
# What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5)
# find the posterior (prior times likelihood)

# First implement a grid approximation with a uniform prior, calculate the posterior and plot the results 
# Then implement a quadratic approximation

# Riccardo 

# grid approximation
dens = 20 
p_grid = seq(from = 0, to = 1, length.out = dens)
# define the prior 
prior = rep(1, dens) # flat (equal possibility for all outcomes)
# prior = ifelse(p_grid < 0.5 , 0 , 1) # unreasonably optimistic prior (informed prior)
# prior <- exp(-5*abs( p_grid - 0.5) #conservative
# Compute the likelihood at each value in grid
R_likelihood <- dbinom( 3 , size=6 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
R_unstd.posterior <- R_likelihood * prior
# Standardize the posterior (so it sums to 1)
R_posterior <- R_unstd.posterior / sum(R_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=R_posterior,prior=prior,likelihood=R_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")
# the probability that Riccardo knows more than chance
sum(R_posterior[p_grid > 0.5])

# quadratic approximation
cogsci_knowledge.qa = map(
  alist(
    c ~ dbinom(6,p) , # binomial likelihood (count of corrects)
    p ~ dunif(0,1)    # uniform prior
  ) ,
  data=list(c=3))

# display summary of quadratic approximation 
precis(cogsci_knowledge.qa)
#   Mean StdDev 5.5% 94.5%
# p  0.5    0.2 0.17  0.83
# the std.dev is the std.dev of the posterior distribution, while the mean value is peak
# Assuming the posterior is Gaussian, it is maximized at 0.5, and its standard deviation is 0.2

```

2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

```{r}
# Kristian

prior = rep(1, dens) # flat (equal possibility for all outcomes)
# prior = ifelse(p_grid < 0.5 , 0 , 1) # unreasonably optimistic prior (informed prior)
# prior <- exp(-5*abs( p_grid - 0.5) #conservative
# Compute the likelihood at each value in grid
K_likelihood <- dbinom( 2 , size=2 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
K_unstd.posterior <- K_likelihood * prior
# Standardize the posterior (so it sums to 1)
K_posterior <- K_unstd.posterior / sum(K_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=K_posterior,prior=prior,likelihood=K_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")
# the probability that Kristian knows more than chance
sum(K_posterior[p_grid > 0.5])

# Josh 

prior = rep(1, dens) # flat (equal possibility for all outcomes)
# prior = ifelse(p_grid < 0.5 , 0 , 1) # unreasonably optimistic prior (informed prior)
# prior <- exp(-5*abs( p_grid - 0.5) #conservative
# Compute the likelihood at each value in grid
J_likelihood <- dbinom( 160 , size=198 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
J_unstd.posterior <- J_likelihood * prior
# Standardize the posterior (so it sums to 1)
J_posterior <- J_unstd.posterior / sum(J_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=J_posterior,prior=prior,likelihood=J_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")
# the probability that Josh knows more than chance
sum(J_posterior[p_grid > 0.5])

# Mikkel

prior = rep(1, dens) # flat (equal possibility for all outcomes)
# prior = ifelse(p_grid < 0.5 , 0 , 1) # unreasonably optimistic prior (informed prior)
# prior <- exp(-5*abs( p_grid - 0.5) #conservative
# Compute the likelihood at each value in grid
M_likelihood <- dbinom( 66 , size=132 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
M_unstd.posterior <- M_likelihood * prior
# Standardize the posterior (so it sums to 1)
M_posterior <- M_unstd.posterior / sum(M_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=M_posterior,prior=prior,likelihood=M_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")
# the probability that Mikkel knows more than chance
sum(M_posterior[p_grid > 0.5])



```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}

# Riccardo (updated prior)

# grid approximation
dens = 20 
p_grid = seq(from = 0, to = 1, length.out = dens)
# define the prior 
prior2 = dnorm(p_grid, mean = 0.8, sd = 0.2)
# Compute the likelihood at each value in grid
R_likelihood <- dbinom( 3 , size=6 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
R_unstd.posterior <- R_likelihood * prior2
# Standardize the posterior (so it sums to 1)
R_posterior <- R_unstd.posterior / sum(R_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=R_posterior,prior=prior2,likelihood=R_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")
# the probability that Riccardo knows more than chance
sum(R_posterior[p_grid > 0.5])

# Kristian (updated prior)

# define the prior 
prior2 = dnorm(p_grid, mean = 0.8, sd = 0.2)
# Compute the likelihood at each value in grid
K_likelihood <- dbinom( 2 , size=2 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
K_unstd.posterior <- K_likelihood * prior2
# Standardize the posterior (so it sums to 1)
K_posterior <- K_unstd.posterior / sum(K_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=K_posterior,prior=prior2,likelihood=K_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")
# the probability that Kristian knows more than chance
sum(K_posterior[p_grid > 0.5])

# Josh (updated prior)

# define the prior 
prior2 = dnorm(p_grid, mean = 0.8, sd = 0.2)
# Compute the likelihood at each value in grid
J_likelihood <- dbinom( 160 , size=198 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
J_unstd.posterior <- J_likelihood * prior2
# Standardize the posterior (so it sums to 1)
J_posterior <- J_unstd.posterior / sum(J_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=J_posterior,prior=prior2,likelihood=J_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")
# the probability that Josh knows more than chance
sum(J_posterior[p_grid > 0.5])

# Mikkel (updated prior)

# define the prior 
prior2 = dnorm(p_grid, mean = 0.8, sd = 0.2)
# Compute the likelihood at each value in grid
M_likelihood <- dbinom( 66 , size=132 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
M_unstd.posterior <- M_likelihood * prior2
# Standardize the posterior (so it sums to 1)
M_posterior <- M_unstd.posterior / sum(M_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=M_posterior,prior=prior2,likelihood=M_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")
# the probability that Mikkel knows more than chance
sum(M_posterior[p_grid > 0.5])
```


4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
# Riccardo (new data + uniform prior + updated prior)

# Compute the likelihood at each value in grid
R_likelihood <- dbinom( 300 , size=600 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by uniform prior)
R_unstd.posterior <- R_likelihood * prior
# Standardize the posterior (so it sums to 1)
R_posterior <- R_unstd.posterior / sum(R_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=R_posterior,prior=prior,likelihood=R_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability") + labs(title = "uniform prior")

# Compute the posterior (likelihood by updated prior)
R_unstd.posterior <- R_likelihood * prior2
# Standardize the posterior (so it sums to 1)
R_posterior <- R_unstd.posterior / sum(R_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=R_posterior,prior=prior2,likelihood=R_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability") + labs(title = "updated prior")

# Kristian (new data + uniform prior + updated prior)

# Compute the likelihood at each value in grid
K_likelihood <- dbinom( 200 , size=200 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by prior)
K_unstd.posterior <- K_likelihood * prior
# Standardize the posterior (so it sums to 1)
K_posterior <- K_unstd.posterior / sum(K_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=K_posterior,prior=prior,likelihood=K_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability") + labs(title = "uniform prior")

# Compute the posterior (likelihood by updated prior)
K_unstd.posterior <- K_likelihood * prior2
# Standardize the posterior (so it sums to 1)
K_posterior <- K_unstd.posterior / sum(K_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=K_posterior,prior=prior2,likelihood=K_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability") + labs(title = "updated prior")

# Josh (new data + uniform prior + updated prior)

# Compute the likelihood at each value in grid
J_likelihood <- dbinom( 1600 , size=1980 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 

# Compute the posterior (likelihood by uniform prior)
J_unstd.posterior <- J_likelihood * prior
# Standardize the posterior (so it sums to 1)
J_posterior <- J_unstd.posterior / sum(J_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=J_posterior,prior=prior,likelihood=J_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability") + labs(title = "uniform prior")

# Compute the posterior (likelihood by updated prior)
J_unstd.posterior <- J_likelihood * prior2
# Standardize the posterior (so it sums to 1)
J_posterior <- J_unstd.posterior / sum(J_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=J_posterior,prior=prior2,likelihood=J_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability") + labs(title = "updated prior")


# Mikkel (new data + uniform prior + updated prior)

# Compute the likelihood at each value in grid
M_likelihood <- dbinom( 660 , size=1320 , prob=p_grid ) # first number how many 1's(corrects), second number how many possible 1's. 
# Compute the posterior (likelihood by uniform prior)
M_unstd.posterior <- M_likelihood * prior
# Standardize the posterior (so it sums to 1)
M_posterior <- M_unstd.posterior / sum(M_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=M_posterior,prior=prior,likelihood=M_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")  + labs(title = "uniform prior")

# Compute the posterior (likelihood by updated prior)
M_unstd.posterior <- M_likelihood * prior2
# Standardize the posterior (so it sums to 1)
M_posterior <- M_unstd.posterior / sum(M_unstd.posterior)
# Draw the plot
Data=data.frame(grid=p_grid,posterior=M_posterior,prior=prior2,likelihood=M_likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of correct answer/knowledge of cogsci")+ ylab("posterior probability")  + labs(title = "updated prior")

```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



NOTES FROM CLASS:
data story - 3 points:
- what is the latent variable of interest?
- how does the latent variable express itself in the observable behaviour
- what is the true structure of the data (e.g. random structure).

before collecting the data/conditioning the data 
- we need to define a prior (our info before data collection) (we know nothing = all values are equally plausible)
- we need to calculate the likelihood of each possible parameter value given the data (condition the model by the data)
- calculate the posterior (prior times lieklihood) 

To calculate the likelihood we can:
- Use Bayes- Theorem ((exact,slow, mostoften not plausible))
- Use numeric solutions (tossing coins) (qudadic appricoxiamation)

Grid approximantion
- define a grid of possible parameter values - 3. 8. 00, 02, 0.0, 

```{r}
# define the grid
dens = 20 
p_grid = seq(from = 0, to = 1, length.out = dens)

# define the prior 
prior = rep(1, dens) # flat
# prior = ifelse(p_grid < 0.5 , 0 , 1) # unreasonably optimistic prior (informed prior)
# prior <- exp(-5*abs( p_grid - 0.5) #conservative

# Compute the likelihood at each value in grid
likelihood <- dbinom( 6 , size=9 , prob=p_grid )

# Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior

# Standardize the posterior (so it sums to 1)
posterior <- unstd.posterior / sum(unstd.posterior)

# Draw the plot
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")

```


first what is the chance we will find the expected results
define r=0-1
binomial answer (only 0 or 1 / right or wrong)


